---
title: "analysis"
author: "Sarah Hvid Andersen"
date: "31/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load('dplyr', 'tidyverse', 'opencage', 'raster', 'rgeos', 'leaflet', 'sf', 'mapboxapi')

# set key in the environment
#Sys.setenv(OPENCAGE_KEY = '')
# set the key interactively if it is missing.
#oc_config('')
```

```{r, echo = FALSE}
Sys.setenv(OPENCAGE_KEY = "a7919189d4134270a7d212e72d9ecd82")
# set the key interactively if it is missing.
oc_config('a7919189d4134270a7d212e72d9ecd82')
```


Loading the raw data:

```{r}
df <- read.csv('data/preprocessed/GYM_alle_p.csv', sep = ';', encoding = 'UTF-8')
address <- read_csv('data/preprocessed/ALL_adresser_p.csv')
```

Ensuring that an address has been gathered for all schools:

```{r}
df$got_address <- ifelse(df$Institutionsnummer %in% address$Institutionsnummer, 1, 0)

missing <- df %>% filter(got_address == 0)

unique(missing$Institution)
unique(missing$Institutionsnummer)
```

Merging the datasets:

```{r}
df <- df %>% 
  rename('Hovedområde' = 'X.U.FEFF.Hovedområde') %>% 
  rename_with( ~ gsub("X", "", .x, fixed = TRUE)) %>% 
  rename_with( ~ gsub(".", "-", .x, fixed = TRUE))

df <- df %>% 
  filter_at(vars(`2012-2013`,`2013-2014`,`2014-2015`, `2015-2016`, `2016-2017`, `2017-2018`, `2018-2019`, `2019-2020`, `2020-2021`, `2021-2022`), any_vars(!is.na(.)))

df3 <- df %>% left_join(address, by = 'Institutionsnummer')
```

Getting the coordinates from the addresses:

```{r}
# concatenating strings to get the full address in one column
df3$location1 <- str_c(df3$Adresse, ', ', df3$Postnummer)
df3$location <- str_c(df3$location1, ' ', df3$By)

locations <- as.data.frame(unique(df3$location))
locations <- locations %>% 
  rename('location' = 'unique(df3$location)')

# geocoding the locations
s <- oc_forward_df(locations, placename = location, countrycode = "DK", limit = 1, no_annotations = TRUE)
GYM_locations <- df3 %>% left_join(s)

# saving the results
write_csv(GYM_locations,"data/preprocessed/GYM_locations.csv")
```

Repeating the process with the VET educations:

```{r}
df <- read.csv('data/preprocessed/EUD_grundforlob1_grupper_p.csv', sep = ';', encoding = 'UTF-8')

df <- df %>% 
  rename('Hovedområde' = 'X.U.FEFF.Hovedområde') %>% 
  rename_with( ~ gsub("X", "", .x, fixed = TRUE)) %>% 
  rename_with( ~ gsub(".", "-", .x, fixed = TRUE))

df <- df %>% 
  filter_at(vars(`2015-2016`, `2016-2017`, `2017-2018`, `2018-2019`, `2019-2020`, `2020-2021`, `2021-2022`), any_vars(!is.na(.)))

df3 <- df %>% left_join(address, by = 'Institutionsnummer')

df3$location1 <- str_c(df3$Adresse, ', ', df3$Postnummer)
df3$location <- str_c(df3$location1, ', ', df3$By)

locations <- as.data.frame(unique(df3$location))
locations <- locations %>% 
  rename('location' = 'unique(df3$location)')

s2 <- oc_forward_df(locations, placename = location, countrycode = "DK", limit =1, no_annotations = TRUE)
EUD_locations <- df3 %>% left_join(s2)

write_csv(EUD_locations,"data/preprocessed/EUD_locations.csv")
```

Reloading the data for further analysis

```{r}
df_gym <- read.csv('data/preprocessed/GYM_locations.csv', encoding = 'UTF-8')
df_eud <- read.csv('data/preprocessed/EUD_locations.csv', encoding = 'UTF-8')
```

specify a year to focus the analysis on. I've chosen 2018/2019 to circumvent any covid related effects. 
```{r}
# selecting the useful columns and rows
df_gym <- df_gym %>% 
  dplyr::select(Hovedområde, Mellemgruppe, Institution.x, Institutionsnummer, Institutions.beliggenhedskommune, oc_formatted, oc_lng, oc_lat, X2018.2019, Kommune, Region) %>% 
  rename(Institution = Institution.x,
         Institution_mun = Institutions.beliggenhedskommune,
         long = oc_lng,
         lat = oc_lat,
         Stud_18 = X2018.2019) %>% 
  mutate(Mellemgruppe = as.factor(Mellemgruppe))

df_gym <- df_gym %>% 
  filter_at(vars(Stud_18), any_vars(!is.na(.)))


# manually registering coordinates for a school missing results
df_gym$lat[df_gym$Institution == 'Prins Henriks Skole, Lycee Francais De Copenhague'] <- 55.673920
df_gym$long[df_gym$Institution == 'Prins Henriks Skole, Lycee Francais De Copenhague'] <- 12.548066
df_gym$oc_formatted[df_gym$Institution == 'Prins Henriks Skole, Lycee Francais De Copenhague'] <- 'Frederiksberg Allé 22A, 1820 Frederiksberg'

```

For the VET educations:
```{r}
# selecting the useful columns and rows
df_eud <- df_eud %>% 
    mutate(Gruppe = Undergruppe) %>% 
  dplyr::select(Hovedområde, Undergruppe, Gruppe, Institution.x, Institutionsnummer, Institutions.beliggenhedskommune, oc_formatted, oc_lng, oc_lat, X2018.2019, Kommune, Region) %>% 
  rename(Institution = Institution.x,
         Institution_mun = Institutions.beliggenhedskommune,
         long = oc_lng,
         lat = oc_lat,
         Stud_18 = X2018.2019)

# creating abreviation (from the English names)
df_eud$Gruppe[df_eud$Gruppe == 'Fødevarer, jordbrug og oplevelser (grundforløb 1)'] <- 'FAH'
df_eud$Gruppe[df_eud$Gruppe == "Kontor, handel og forretningsservice (grundforløb 1)"] <- 'ACB' 
df_eud$Gruppe[df_eud$Gruppe == "Omsorg, sundhed og pædagogik (grundforløb 1)"] <- 'CHP'
df_eud$Gruppe[df_eud$Gruppe == "Teknologi, byggeri og transport (grundforløb 1)"] <- 'TCT'

# removing na's
df_eud <- df_eud %>% 
  filter_at(vars(Stud_18), any_vars(!is.na(.)))
```

getting municipality polygons:
get distance from municipality (coord) to the nearest edu group

```{r}
# downloading municipality data
municipalities <- getData("GADM", country = "DNK", level = 2)

municipalities_sf <- st_transform(st_as_sf(municipalities), crs = 4326)

municipalities_sf <- municipalities_sf %>% 
  mutate(NAME_2 = gsub("Århus", "Aarhus", NAME_2)) %>% 
  mutate(NAME_2 = gsub("Høje Taastrup", "Høje-Taastrup", NAME_2)) %>% 
  mutate(NAME_2 = gsub("Vesthimmerland", "Vesthimmerlands", NAME_2))

# paste
municipalities_sf$location <- paste(municipalities_sf$NAME_2, municipalities_sf$ENGTYPE_2, sep = ' ')

# geocode
mun_coords <- oc_forward_df(municipalities_sf,
                           placename = location, countrycode = "DK", limit =1, no_annotations = TRUE)

# correcting coordinates
mun_coords$oc_lat[mun_coords$NAME_2 == 'Bornholm'] <- 55.123448
mun_coords$oc_lng[mun_coords$NAME_2 == 'Bornholm'] <- 14.909327

mun_coords$oc_lat[mun_coords$NAME_2 == 'Fanø'] <- 55.415258
mun_coords$oc_lng[mun_coords$NAME_2 == 'Fanø'] <- 8.403257

mun_coords$oc_lat[mun_coords$NAME_2 == 'Kerteminde'] <- 55.460983
mun_coords$oc_lng[mun_coords$NAME_2 == 'Kerteminde'] <- 10.644460

mun_coords$oc_lat[mun_coords$NAME_2 == 'Odense'] <- 55.401549
mun_coords$oc_lng[mun_coords$NAME_2 == 'Odense'] <- 10.374097

mun_coords$oc_lat[mun_coords$NAME_2 == 'Lejre'] <- 55.632084
mun_coords$oc_lng[mun_coords$NAME_2 == 'Lejre'] <- 11.874893
 
# saving and rereading the results

st_write(mun_coords, "data/preprocessed/mun_coord.shp")
mun_coords = st_read("data/preprocessed/mun_coord.shp")

mun_coords_df <- mun_coords %>% 
  dplyr::select(NAME_2, oc_lat, oc_lng) %>% 
  rename(Kommune = NAME_2)

mun_coords_df <- as.data.frame(mun_coords_df) %>% 
  dplyr::select(-c(geometry))

```

Getting the distance from a municipality and to the nearest school of each group type (hf, stx, hhx, htx)
Creating a dataframe with school coordinates and the municipality of pupil admission coordinates (kommune column) and getting the distance for each of these. 

```{r}
# create merge by kommune for distance calculation
df_gym <- df_gym %>% left_join(mun_coords_df, by = 'Kommune')
df_eud <- df_eud %>% left_join(mun_coords_df, by = 'Kommune')

# for the big dataframe, need to drop nas, where the living mun was unspecified
df_gym <- df_gym %>% drop_na(oc_lng)
df_eud <- df_eud %>% drop_na(oc_lng)

# Takes 3.1 minute for gym, 1.2 for eud. times taken at 13 weekend, no rush hour
get_distance <- function(df) {
  var_list = list()
  dist_list = list()
  
  start_time <- Sys.time()
  for (row in 1:nrow(df)) {
      school_long <- df[row, "long"]
      school_lat  <- df[row, "lat"]
      school_name <- df[row, "Institution"]
      
      kom_long <- df[row, "oc_lng"]
      kom_lat  <- df[row, "oc_lat"]
      kom_name <- df[row, "Kommune"]
      
      cat(school_name,'------>', kom_name, '\n')
      print(row +1)
  
      route <- mb_directions(
        origin = c(school_long, school_lat),
        destination = c(kom_long, kom_lat),
        profile = "driving") 
      
      dist_list[[row]] <- route
      var_list[[row]] <- c(school_name, kom_name)
    }
  end_time <- Sys.time()
  print(end_time - start_time)
  
  results <- list("variables" = var_list, "distances" = dist_list)
  return(results)
}

# apply function to gym and eud dataframes
result_eud <- get_distance(df_eud)
result_gym <- get_distance(df_gym)

# gather the results into one SF dataframe
var_df_eud <- do.call(rbind, result_eud$variables) %>%
  as.data.frame()
dist_sf_eud <- dplyr::bind_rows(result_eud$distances)
dist_sf_eud <- cbind(dist_sf_eud, var_df_eud)

var_df_gym <- do.call(rbind, result_gym$variables) %>%
  as.data.frame()
dist_sf_gym <- dplyr::bind_rows(result_gym$distances)
dist_sf_gym <- cbind(dist_sf_gym, var_df_gym)

# save copy of distances including geometry shape of the paths chosen
st_write(dist_sf_eud, "data/preprocessed/dist_eud_mun.shp")
st_write(dist_sf_gym, "data/preprocessed/dist_gym_mun.shp")

# append to existing dataframe
df2 <- as.data.frame(dist_sf_gym) %>% 
  dplyr::select(-geometry)
df_gym <- cbind(df_gym, df2)

df2 <- as.data.frame(dist_sf_eud) %>% 
  dplyr::select(-geometry)
df_eud <- cbind(df_eud, df2)

## create gathered df
df_gym <- df_gym %>% 
  mutate(Undergruppe = NA) %>% 
  rename(Gruppe = Mellemgruppe)

df <- full_join(df_gym, df_eud)
```

Getting some descriptive statistics on the final data.

```{r}
# get municipality student sum
Mun_summary <- df %>% 
  group_by(Kommune) %>% 
  summarise(Mun_sum = sum(Stud_18))

df <- df %>% left_join(Mun_summary, by = 'Kommune')

df <- df %>% mutate(Mun_percent = round((Stud_18 / Mun_sum), digits = 3))

# percent of municipality that chose which main area
df$Mun_main <- paste(df$Kommune, df$Hovedområde)

Main_summary <- df %>% 
  group_by(Mun_main) %>% 
  summarise(Mun_main_sum = sum(Stud_18))

df <- df %>% left_join(Main_summary, by = 'Mun_main')

df <- df %>% mutate(Mun_main_percent = round((Mun_main_sum / Mun_sum), digits = 3))

# percent of municipality that chose which group
df$Mun_group <- paste(df$Kommune, df$Gruppe)

Group_summary <- df %>% 
  group_by(Mun_group) %>% 
  summarise(Mun_group_sum = sum(Stud_18))

df <- df %>% left_join(Group_summary, by = 'Mun_group')

df <- df %>% mutate(Mun_group_percent = round((Mun_group_sum / Mun_sum), digits = 3))

# save the results
write_csv(df, 'data/preprocessed/all_results.csv')
df <- read_csv('data/preprocessed/all_results.csv')

# creating a distinct education data frame for the interactive map
df_edu <- distinct(df, Institution, Gruppe, .keep_all = TRUE)

# Find a center point for each municipality for the interactive map
centers <- data.frame(gCentroid(municipalities, byid = TRUE))
centers$mun <- municipalities$NAME_2
```

What is the minimum travel time to each education, by municipality?
```{r}
# creating a df that only keeps the min distance values for each municipality and group
df_min <- df %>% 
  group_by(Kommune, Gruppe)%>% 
  slice(which.min(duration))

df_min <- df_min %>% 
  mutate(Gruppe = factor(Gruppe, levels = c('CHP', 'ACB', 'FAH', 'TCT', 'Stx', 'Hf', 'Hhx', 'Htx', 'Øvrige')))

# min distance for each municipality and main area
df_main <- df %>% 
  group_by(Kommune, Hovedområde)%>% 
  slice(which.min(duration))

df_main <- df_main %>% 
  mutate(Gruppe = factor(Gruppe, levels = c('CHP', 'ACB', 'FAH', 'TCT', 'Stx', 'Hf', 'Hhx', 'Htx', 'Øvrige'))) %>% 
  mutate(Education = factor(Hovedområde, levels = c('Gymnasiale uddannelser', 'Erhvervsfaglige uddannelser')))


# Difference in distance by main area and group
df <- df %>% 
  mutate(Gruppe = factor(Gruppe, levels = c('CHP', 'ACB', 'FAH', 'TCT', 'Stx', 'Hf', 'Hhx', 'Htx', 'Øvrige')))

df %>% group_by(Hovedområde) %>% 
  summarise(mean(distance))

df %>% group_by(Gruppe) %>% 
  summarise(mean(distance))
```


Creating interactive map with Leaflet():

```{r}
# creating a separate dataset for each layer in the interactive plot
stx = df_edu %>% filter(Gruppe == 'Stx')
hf = df_edu %>% filter(Gruppe == 'Hf')
htx = df_edu %>% filter(Gruppe == 'Htx')
hhx = df_edu %>% filter(Gruppe == 'Hhx')
chp = df_edu %>% filter(Gruppe == 'CHP')
acb = df_edu %>% filter(Gruppe == 'ACB')
fah = df_edu %>% filter(Gruppe == 'FAH')
tct = df_edu %>% filter(Gruppe == 'TCT')


# using mapbox tiles as this is a license requirement if i plot any of the distances
map <- leaflet() %>%
  addMapboxTiles(style_id = "light-v10", #alternative, streets-v11: The core Mapbox basemap
                 username = "mapbox") 

map %>%  
  addPolylines(data = municipalities_sf,
    stroke=TRUE, 
    fillOpacity = 1,
    color = 'black',
    weight=0.4) %>% 
  
  addLabelOnlyMarkers(data = centers,
                    lng = ~x, lat = ~y, label = ~mun,
                    labelOptions = labelOptions(noHide = TRUE, direction = 'top',
                                                textOnly = TRUE,
                                                textsize = '7.5px')) %>%
  # stx
  addCircleMarkers(stx$long,
                   stx$lat,
                   radius = 4,
                   color = 'turquoise',
                   group = 'STX',
                   popup = paste0(stx$Institution, "<br/>Type: ", stx$Gruppe)) %>%
  # hf
  addCircleMarkers(hf$long,
                   hf$lat,
                   radius = 4,
                   color = 'blue',
                   group = 'HF',
                   popup = paste0(hf$Institution, "<br/>Type: ", hf$Gruppe)) %>%
    # hhx
  addCircleMarkers(hhx$long,
                   hhx$lat,
                   radius = 4,
                   color = 'purple',
                   group = 'HHX',
                   popup = paste0(hhx$Institution, "<br/>Type: ", hhx$Gruppe)) %>%
    # htx
  addCircleMarkers(htx$long,
                   htx$lat,
                   radius = 4,
                   color = 'pink',
                   group = 'HTX',
                   popup = paste0(htx$Institution, "<br/>Type: ", htx$Gruppe)) %>%
    # CHP
  addCircleMarkers(chp$long,
                   chp$lat,
                   radius = 4,
                   color = 'red',
                   group = 'CHP',
                   popup = paste0(chp$Institution, "<br/>Type: ", chp$Gruppe)) %>%
    # acb
  addCircleMarkers(acb$long,
                   acb$lat,
                   radius = 4,
                   color = 'orange',
                   group = 'ACB',
                   popup = paste0(acb$Institution, "<br/>Type: ", acb$Gruppe)) %>%
    # fah
  addCircleMarkers(fah$long,
                   fah$lat,
                   radius = 4,
                   color = 'yellow',
                   group = 'FAH',
                   popup = paste0(fah$Institution, "<br/>Type: ", fah$Gruppe)) %>%
    # tct
  addCircleMarkers(tct$long,
                   tct$lat,
                   radius = 4,
                   color = 'green',
                   group = 'TCT',
                   popup = paste0(tct$Institution, "<br/>Type: ", tct$Gruppe)) %>%

  
  # add layer control
  addLayersControl(overlayGroups = c('STX', 'HF', 'HHX', 'HTX', 'CHP', 'ACB', 'FAH', 'TCT'),
                   options = layersControlOptions(collapsed = FALSE))

```

identify the municipalities that have difficulty in accessing all 8 types:

```{r}
df_access <- df_min %>% 
  filter(distance >= 50, Gruppe != 'Øvrige') %>% 
  mutate(Kommune = factor(Kommune),
         distance = round(distance)) %>% 
  dplyr::select(Region, Kommune, Gruppe, Hovedområde, distance, duration)

write_csv(df_access, 'data/preprocessed/bad_access_muns.csv')
```


Plotting results:

```{r}
# plotting with the percentage of the municipality that chose which group
plot_groups <- function(df) {
ggplot(data = NULL, aes(group = Gruppe)) +
  geom_bar(data=df, 
           aes(fill=Gruppe, y=distance, x=Kommune), position="dodge", stat="identity") +
  geom_point(data=df, 
             aes(y=Mun_group_percent*25, x=Kommune, group=Gruppe), position = position_dodge(width = 0.9), shape=3, size=4, show.legend=FALSE) + 
    xlab('Municipality') + ylab('Distance (km)') +
  ggtitle('The minimum distance to each education group from each municipality')
}

# fitting plot function by region
df_min %>% 
  # esxcluding Læsø from the plot as it severely distorts the plot (outlier)
  filter(Region == 'Region Nordjylland', Kommune != 'Læsø', Gruppe != 'Øvrige') %>%
  plot_groups()

df_min %>% 
  filter(Region == 'Region Midtjylland', Gruppe != 'Øvrige') %>% 
  plot_groups()

df_min %>% 
  filter(Region == 'Region Syddanmark', Gruppe != 'Øvrige') %>% 
  plot_groups()

df_min %>% 
  filter(Region == 'Region Sjælland', Gruppe != 'Øvrige') %>% 
  plot_groups()

df_min %>% 
  filter(Region == 'Region Hovedstaden', Gruppe != 'Øvrige') %>% 
  plot_groups()


# plot for the main group
plot_main <- function(df) {
ggplot(data = NULL, aes(group = Education)) +
  geom_bar(data=df, 
           aes(fill=Education, y=distance, x=Kommune), position="dodge", stat="identity") +
  geom_point(data=df, 
             aes(y=Mun_main_percent*15, x=Kommune, group=Education), position = position_dodge(width = 0.9), shape=3, size=4, show.legend=FALSE) +
  labs(title = "The minimum distance to each education type from each municipality", x = "Municipality", y = "Distance (km)", color = "Education\n")+
    scale_fill_discrete(labels = c("GE", "VET"))
}

# plotting by region

df_main %>% 
  filter(Region == 'Region Nordjylland', Kommune != 'Læsø') %>% 
  plot_main()

df_main %>% 
  filter(Region == 'Region Midtjylland') %>% 
  plot_main()

df_main %>% 
  filter(Region == 'Region Syddanmark') %>% 
  plot_main()

df_main %>% 
  filter(Region == 'Region Sjælland') %>% 
  plot_main()

df_main %>% 
  filter(Region == 'Region Hovedstaden') %>% 
  plot_main()
```

